---
title: "WUP_04"
author: "Taylor Stohl"
date: "6-11-2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

```{r}
gssdata=read_csv("GSScsv.csv")
```
# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(gssdata,aes(x=DEGREE,y=WORDSUM))
```
## Now add in boxplot

```{r}
basicCN + geom_boxplot()
```
  
It appears that if we run this test, the LT HIGHSCHOOL is drastically different than JUNIOR COLLEGE and LT HIGHSCHOOL is also drastically different than GRADUATE. Other differences aren't as obvious. If the Box for LT HIGHSCHOOL was moved over to the left, it wouldn't touch the JUNIOR COLLEGE box or the GRADUATE box at all, meaning that these two categories have very different results in WORDSUM test scores.




### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


```{r}
favstats(WORDSUM~DEGREE,data=gssdata)
```

It appears that the BACHELOR(9.2), GRADUATE(10.09), and JUNIOR COLLEGE(9.91)  categories have higher means. However, LT HIGHSCHOOL(7.75)and HIGHSCHOOL(8.28) have a lower mean by a couple points.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
model1 <- aov(WORDSUM~DEGREE,data=gssdata)
summary.aov(model1)
```

The P-value is 0.000236. There is a three star difference which means that there is a very significant difference in WORDSUM test scores and DEGREE. The chances of getting these results is 1 in 23000. The null hypothesis is rejected. However, we still need clarification on which categories had different results.


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 13.2, 2, 2))
plot(TukeyHSD(model1),las=2)
```

Of all the possible differences, HIGHSCHOOL and GRADUATE, and LT HIGHSCHOOL and GRADUATE are different in value because the line comparing these values doesn't pass though the dotted line(0). If the line passes through 0, that means there is no difference.


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

 Yes, there is a difference in WORDSUM test scores an DEGREE. HIGHSCHOOL and GRADUATE and LT HIGHSCHOOL and GRADUATE had different results in the WORDSUM test. The null hupothesis that the there is no difference between the average score on the WORDSUM test and level of schooling DEGREE is rejected. 
